---
title: "ANOVA Practice"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, warning = FALSE, message = FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(emmeans)
library(multcomp)
```

## Questions {.tabset}

### As we have done before:

1.  Make a new project folder (directory) containing a .Rproj file. Use
    your last name as the first part of the file name (e.g.
    "Mylastname_ANOVA_practice").

2.  Open a blank script file (.R) and save it in your project folder

3.  Copy any data files you need for answering questions into the
    project folder.

4.  Load necessary libraries

5.  Make each answer an object (like we did in the previous practices,
    "quest1 \<- code...code...code")

6.  Save each figure as a .png file.

7.  Provide a results statement in the format expected in a journal
    article.

------------------------------------------------------------------------

```{r import data, include=FALSE}
medley <- read.csv("medley.csv")
#diseases <- readr::read_csv("diseases.csv")
```

### Practice 1

Medley and Clements (1998) sampled a number of stations (between four
and seven) on six streams known to be polluted by heavy metals in the
Rocky Mountain Region of Colorado. They recorded zinc concentration,
species diversity of the diatom community, and the name of each stream.

`medley.csv` file contents:

**stream** - name of streams in the Rocky Mountain region of Colorado,
USA

**zinc_level** - categorical zinc concentration level (HIGH=high,
MED=medium, LOW=low, BACK=background)

**diversity** - Shannon-Wiener species diversity of diatoms

**reference_number** - alternative categorical zinc concentration level
(1=background, 2=low, 3=medium, 4=high)

Using the `medley.csv` data set,

1.  Determine whether the mean diatom species diversity differs across
    the four zinc-level groups

2.  Determine whether different levels of zinc pollution differ in their
    effect on diatom species diversity

3.  Present a bar chart figure depicting means and standard errors for
    the four levels of pollution.

4.  Arrange the bars to proceed in ascending order of pollution (B \< L
    \< M \< H). Hint, after importing the data, immediately change the
    order of factor levels (the default is alphabetical, but that
    doesn't make sense here).

```{r quest 1 relevel}
#Reorder the factor levels to be more intuitive.
medley$zinc_level <- factor(medley$zinc_level, levels = c('BACK', 'LOW', 'MED', 'HIGH'))
#check to make sure they were reordered
unique(medley$zinc_level)
```

Start by visualizing the data, easiest methods are a point plot and a
box plot (doesn't require you to summarize data first).

```{r quest 1 boxplot, echo=FALSE}
ggplot(medley, aes(x = zinc_level, y = diversity)) +   # data & aesthetic mapping
  geom_point()

ggplot(medley, aes(x = zinc_level, y = diversity)) +   # data & aesthetic mapping
  geom_boxplot()
```

These figures don't actually help visualize differences all that much,
especially because there are some potential outliers that are tricking
our eye and no error bars to help visually determine pairwise group
differences. Lets summarize the data so we can display the means and
error bars. If your levels of zinc are in a different order (such as
alphabetical), then you didn't properly re-level the categorical
`zinc_level` variable.

```{r quest 1 stats, echo=FALSE}
stats_medley <- medley %>%  # names of the new data frame and the data frame to be summarized
  group_by(zinc_level) %>%  # grouping variable
  summarise(mean = mean(diversity),  # mean of each group
            n    = n(),        # sample size per group
            sd   = sd(diversity),    # standard deviation of each group
            se   = sd(diversity) / sqrt(n()))  # standard error of each group
stats_medley
```

Now that we have summary statistics for the `medley` data set, we can
make a more informative figure.

```{r quest 1 figure, echo=FALSE}
p1 <- ggplot(stats_medley, aes(x = zinc_level, y = mean)) +   # data & aesthetic mapping
  geom_bar(stat = "identity") +                               # bars represent means
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2)  # error bars
p1
```

That is much more helpful because we can clearly see both the means and
standard errors to infer pairwise differences among stream types. The
HIGH concentration zinc streams clearly have lower species diversity
than the others. First we need to fit a linear model to see whether
there are *overall* differences among stream types.

```{r quest 1 model, echo=FALSE}
model_zinc <- lm(diversity ~ zinc_level, data=medley)
summary(model_zinc)
```

From the ANOVA report at the bottom, there appears to be overall
differences (p = 0.017) in diversity among the streams with different
concentrations of zinc. This indicates that *at least one stream* is
significantly different from the others in species diversity (reject
null hypothesis that all stream types have the same diversity).

Now we use the `emmeans` package to make pairwise comparisons among the
four stream types.

```{r quest 1 calc emmeans, echo=FALSE}
#Multiple Contrasts using EMmeans package
emm_model_zinc<- emmeans(model_zinc, "zinc_level")
pairs(emm_model_zinc)
```

It looks like the only significant difference in species diversity was
between the LOW and HIGH zinc streams (p = 0.011). Lets do a quick plot
to visualize this difference. Yep, no overlap between error bars of LOW
and HIGH groups.

```{r quest 1 plot emmeans, echo=FALSE}
#visually compare marginal means (LS means)
plot(emm_model_zinc, comparisons = TRUE) +
  coord_flip()
```

Now we can try to describe all pairwise comparisons among groups
ourselves, or we can use a Connecting Letters Differences (CLD) report
to help.

```{r quest 1 multcomp CLD, echo=FALSE}
#connecting letters differences report
multcomp::cld(emm_model_zinc, Letters=letters)
```

Transcribe the group letters onto the bar chart you made previously. Be
careful, the `multcomp` package does not respect the reordering of
levels you did previously; it sorts groups by lowest to highest mean.
Transcribe letters carefully.

(*Sidenote: Can you think of a way of forcing R to sort the CLD report
by your re-leveled `zinc_level` variable rather than by the default of
increasing `emmean`?*)

If you previously made the chart an object (like you should always do)
then it will be easy to call up here and add text annotations or text
boxes depicting differences among streams.

```{r quest 1 add CLD}
p1 +
  annotate("text", x=4, y=1.5, label="a") +    #HIGH
  annotate("text", x=3, y=2, label="a,b") +    #MED
  annotate("text", x=1, y=2.25, label="a,b") + #BACK
  annotate("text", x=2, y=2.25, label="b") +  #LOW
  #make plot pretty
  xlab("Zinc Level") +
  ylab("Diatom Species Diversity") +
  theme_bw()
```

Results Statement goes here.....

------------------------------------------------------------------------

### Practice 2

Using the `ChickWeight` data set built into R, test the hypothesis that
diet type influences growth (and therefore weight) of chicks. Call the
`ChickWeight` data set and notice that the data are a time-series;
multiple measurements of each chick taken across different time periods.
Because these data are therefore not independent (one of our
assumptions) it will be necessary to reduce the data set to one that is.
Use only the final measurements for each chick (n=50 chicks at Time=21)
and discard the others (we then have one measurement of each chick).

```{r}
ChickWeight <- ChickWeight %>%
  filter(Time == 21)

str(ChickWeight)
```

Typing `?ChickWeight` gives more information on the data. Run the
example code provided in the help file if you want to see the growth
trajectories.

1.  Are there differences among diets **overall**?
2.  Are there **pairwise differences** between the different diets?
3.  Present a point chart (`geom_point`) showing means and standard
    errors for the four levels of pollution.
4.  Use Connecting Letters Differences (CLD) notation to show how diet
    groups differ.

Make summary statistics of the data set.

```{r quest 2 stats, echo=FALSE}
stats_chicks <- ChickWeight %>%  # names of the new data frame and the data frame to be summarized
  group_by(Diet) %>%  # grouping variable
  summarise(mean = mean(weight),  # mean of each group
            n    = n(),        # sample size per group
            sd   = sd(weight),    # standard deviation of each group
            se   = sd(weight) / sqrt(n()))  # standard error of each group
stats_chicks
```

```{r quest 2 figure, echo=FALSE}
p2 <- ggplot(stats_chicks, aes(x = Diet, y = mean)) +   # data & aesthetic mapping
  geom_point(size=4) +                               # bars represent means
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2)  # error bars
p2
```

```{r quest 2 model, echo=FALSE}
model_chicks <- lm(weight ~ Diet, data=ChickWeight)
summary(model_chicks)
```

```{r quest 2 emmeans, echo=FALSE}
#Multiple Contrasts using EMmeans package
emm_model_chicks<- emmeans(model_chicks, "Diet")
pairs(emm_model_chicks)
```

```{r quest 2 plot emmeans, echo=FALSE}
#visually compare marginal means (LS means)
plot(emm_model_chicks, comparisons = TRUE) +
  coord_flip()
```

```{r quest 2 multcomp CLD, echo=FALSE}
#connecting letters differences report
multcomp::cld(emm_model_chicks, Letters=letters)
```

Make the plot annotated with CLD letters and pretty it up.

```{r quest 2 add CLD}
p2 +
  annotate("text", x=1, y=225, label="a")+
  annotate("text", x=2, y=275, label="a,b")+
  annotate("text", x=4, y=290, label="b")+
  annotate("text", x=3, y=305, label="b") +
  ylab("Final Chicken Weight, grams (mean +/- 1 SE)")+
  theme_bw()
```

Results Statement goes here.....

------------------------------------------------------------------------

### Practice 3

Basil is one of the world most important commercial herbs. Like many
herbaceous species, basil has several desirable compounds used by
humans. Basildiol is the active ingredient in basil that gives the
distinct bite and aroma. You are working for a commercial lab interested
in extracting the maximum basildiol and you have three candidate
strains.

Is there an **overall** difference among strains?

Are there **pairwise** differences between the strains?

```{r quest 3 import data, include=FALSE}
basil <- read.csv("basil.csv")
```

```{r quest 3 stats, echo=FALSE}
stats_basil <- basil %>%  # names of the new data frame and the data frame to be summarized
  group_by(strain) %>%  # grouping variable
  summarise(mean = mean(basildiol),  # mean of each group
            n    = n(),        # sample size per group
            sd   = sd(basildiol),    # standard deviation of each group
            se   = sd(basildiol) / sqrt(n()))  # standard error of each group
stats_basil
```

```{r quest 3 figure, echo=FALSE}
p3 <- ggplot(stats_basil, aes(x = strain, y = mean)) +   # data & aesthetic mapping
  geom_bar(stat = "identity") +                               # bars represent means
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2)  # error bars
p3
```

```{r quest 3 model, echo=FALSE}
model_basil <- lm(basildiol ~ strain, data=basil)
summary(model_basil)
```

```{r quest 3 calc emmeans, echo=FALSE}
#Multiple Contrasts using EMmeans package
emm_model_basil<- emmeans(model_basil, "strain")
pairs(emm_model_basil)
```

```{r quest 3 multcomp CLD, echo=FALSE}
#connecting letters differences report
multcomp::cld(emm_model_basil, Letters=letters)
```

```{r quest 3 add CLD}
p3 +
  annotate("text", x=1, y=15, label="a") +      #dank
  annotate("text", x=3, y=17.5, label="a") +    #dream
  annotate("text", x=2, y=22.5, label="b") +      #diesel
  #make plot pretty
  xlab("Basil Strain") +
  ylab("Concentration of Basildiol (% Dry Mass)") +
  theme_bw(base_size=15) +
  annotate("text", x=1.35, y=25, label="F2,26 = 11.06, P = 0.0003, R^2 = 0.41")+
  aes(fill=strain) +
  theme(legend.position="none")
```

Results Statement goes here.....

------------------------------------------------------------------------
